
\title{\vspace{-3cm}
      Expecting to get it: \\ An endowment effect for information
      }

% ADD LINK TO UPTODATE VERSION AND SUPPLEMENTARY+REPLICATION MATERIAL

\author{Tabaré Capitán
          \thanks{Corresponding author: \href{mailto:Tabare.Capitan@gmail.com}{\texttt{Tabare.Capitan@gmail.com}}. For their useful comments or thoughtful discussions, we thank Hunt Allcott, Alec Brandon, James Heckman, Stephan Kroll, John List, Daniel Millinet, and Stephen Newbold. We also thank participants in seminars at UA Anchorage, the University of Chicago, Colorado State University, the 2019 North American Economic Science Association meetings, and the 2020 Bogotá Experimental Economics Conference. Financial support from USDA/NIFA (grant number 2016-09907) is gratefully acknowledged. The collection of data for this study was approved by the IRB at University of Wyoming.}
        \and
        Linda Thunström
        \and
        Klaas van ‘t Veld
          \\ \small{Department of Economics, University of Wyoming}
        \and
        Jonas Nordström
          \\ \small{Lund University and University of Copenhagen}
        }

\maketitle

\thispagestyle{empty}   % Suppress page number (must be under \maketitle)

\begin{abstract}

\noindent
We show, both theoretically and empirically, that an endowment effect for information arises from the expectation of getting information. Furthermore, we show that this endowment effect matters to welfare analysis of information policies, and can be put to advantage in situations where widespread dissemination of information (e.g., about health or environmental risks) is socially desirable.
\\
\\
\textit{Keywords:} expectations, reference-dependent preferences, preferences for information, information avoidance.
\\
\textit{JEL classification:} D01, D80, D83, D84, D91, C91.

\end{abstract}

\clearpage

\pagenumbering{arabic} % Start numbering in page 2 from #1
